%------------------------------------------------------
% QUOTE
% If you really feel like adding a quote
% to your page, uncomment out the following.
%------------------------------------------------------
%\begin{savequote}[45mm]
%When theory and experiment agree, 
%that is the time to be especially suspicious. 
%\qauthor{Niels Bohr}
%\end{savequote}
%------------------------------------------------------
% QUOTE
% If you really feel like adding a quote
% to your page, uncomment out the previous.
%------------------------------------------------------


\chapter{Conclusion}

This thesis explored some potential advantages offered by Approximate Bayesian Computation (ABC) in the context of geophysical inversions. Chapter \ref{SGE} illustrated how ABC can use the whole wealth  of information available by opening the likelihood to drive a more diagnostic inversion scheme. This diagnostic inversion improves in comparison to traditional likelihood machinery by ensuring each model update has impact by providing information on where the model is inadequate and how it can improve to match the observed data. A specific result is rapid convergence to low misfit models while retaining the same formal statistical guarantees offered by probabilistic formulations of geophysical inverse problems. Algorithms which rapidly find and explore low misfit models are necessary if we wish to push the limits in the resolution and scale of solvable geophysical inverse problems which fundamentally require a probabilistic approach. This thesis represents the first attempt at using ABC in joint geophysical inversions, providing an initial connection to the rapidly expanding field of likelihood-free methods. This value is highlighted by the new likelihood-free methods consistently emerging which may benefit geophysics \citep{papamakarios2016fast,song2017nice}.\par

While the results presented in this exploratory analysis are overall positive, more comprehensive studies are required to fully uncover the potential of ABC in geophysics. Firstly, the range of tests for the diagnostic ABC inversion needs to be expanded to larger and more realistic Earth models. The intrinsic trade-off between tolerance, sampler efficiency and accuracy of the posterior PDF needs to be addressed in more detail if ABC is to become a robust practical option. Further exploration is also needed in how to best drive a diagnostic scheme given the available information, while still retaining formal statistical guarantees. The potential for comparative improvement is not restricted to this angle. It is worthwhile assessing the impact of simplifying assumptions about the nature of modelization and data uncertainty, compared with its true nature. There is also scope for assessing the model adequacy given the data, as opposed to relative adequacy against other models. \par

This thesis demonstrates how ABC can be used to build upon the progress made in probabilistic tomography.
